\vspace{-1.5em} \bf A general and efficient representation of ancestral recombination graphs

### 2023-10-20 paper.tex
Fill in Example inferred ARGs section

\documentclass{article}
\usepackage[letterpaper,top=2cm,bottom=2.5cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{geometry}%
\usepackage{color}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{nameref}
\usepackage[right]{lineno}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{math,calc,positioning}
\usepackage{url}
\usepackage[symbol]{footmisc}
% JK: turning this off for the moment as I keep clicking through on links
% to the bibliography while reading the text and it's intensely annoying.
% Can reinstate when we're ready to preprint
% \usepackage[hidelinks]{hyperref}

@@

\linenumbers
\section{Introduction}
\label{sec-intro}
Estimating the genetic genealogy of a set of genome sequences
under the influence of recombination,
usually known as an Ancestral Recombination Graph (ARG), is a long-standing
goal in genetics.
Broadly speaking, an ARG describes the different paths of genetic inheritance
caused by recombination, and encodes a sequence of correlated genealogical
trees along the genome.
Recent breakthroughs
in large-scale inference
methods~\citep{rasmussen2014genome,kelleher2019inferring,speidel2019method,
schaefer2021ancestral,wohns2022unified,zhang2023biobank,zhan2023towards}
and data representation~\citep{kelleher2016efficient,kelleher2018efficient}
have raised the realistic prospect of ARG-based analysis becoming a standard part
of the population and statistical genetics toolkit~\citep{hejase2020summary}.
Applications using inferred ARGs as input have begun to
appear~\citep{osmond2021estimating,fan2022genealogical,hejase2022deep,zhang2023biobank,
nowbandegani2023extremely,ignatieva2023distribution,fan2023likelihood}
and many more are sure to
follow~\citep{harris2019database,harris2023using}.
See \citet{lewanski2023era} for a biologically oriented introduction to ARGs.

@@

% JK: this para need a major update, it's quite weak right now.
The word ``ARG'' is now widely used to refer to genetic genealogies in
general~\citep[e.g.][]{mathieson2020ancestry,hejase2020summary,
schaefer2021ancestral,harris2023using,zhang2023biobank,fan2023likelihood},
encompassing the varied outputs from modern simulation and
inference methods~\citep{rasmussen2014genome, palamara2016argon, haller2018tree,
kelleher2019inferring, speidel2019method, baumdicker2021efficient, zhang2023biobank}.
This broad usage means there is a lack of consensus on the
exact object being defined, which causes problems, for example,
when evaluating inferred ARGs, and even when defining
what the criteria for success should be.
It also means there is no shared format for interchanging ARG data between
programs, resulting in the assorted negative consequences with which
geneticists are only too familiar~\citep{excoffier2006computer,gower2022demes}.
Perhaps most importantly, definitions linked to
the original stochastic process lead to an
overly narrow interpretation of ARGs, which is significantly
out of step with modern methods and large-scale datasets, and
can lead to missed opportunities in their application and usage.
Such problems go much deeper than pedantic discussion over terminology;
they directly inform what we should be attempting to estimate, how we
assess success, and the expected limitations on the precision
and accuracy of our inferences.
% Thanks for the phrase Gideon and Alex!
If not addressed, these problems will
impede progress in applying genetic genealogies to understand the biological world.

@@

\begin{figure} \begin{center}
\includegraphics[width=\textwidth]{illustrations/inference.pdf} \end{center}
\caption{\label{fig-inferred-args} Inference results for
(A) \kwarg, (B) \argweaver, (C) \tsinfer,
and (D) \relate\
for 11 samples of a 2.7kb region of the \textit{Drosophila melanogaster} ADH
locus~\citep{kreitman1983nucleotide}.
The \argweaver\ example
is a sample from the inferred posterior distribution.
Edge colours are based on the time of the edge's child node
(lighter: older; darker: younger).
Position on the X and Y axes for the graph visualization is arbitrary.
Line width and node colour are as described in Fig.\ref{fig-simplification}.
The genomic intervals associated with edges are shown below
each ARG, where each row is an edge
and the X axis reflects genomic position.
Vertical lines indicate breakpoints between local trees.
}
\end{figure}

% TODO add explanation of the bottom row.
% https://github.com/tskit-dev/what-is-an-arg-paper/issues/386
The example of successive simplification of a simulated ARG
in Fig.~\ref{fig-simplification} illustrates the point
that is is possible to \emph{store} arbitrary levels of precision
about recombination in a gARG.
It is also possible to \emph{infer} various levels of precision about
recombination.
To illustrate this, and to demonstrate the structural
heterogeneity of the ARGs inferred  by current methods,
Fig.~\ref{fig-inferred-args} shows the results of ARG inference
% To illustrate these points and to emphasise the heterogeneity
% of the ARGs inferred by current methods, we inferred ARGs for the
for the classical \citet{kreitman1983nucleotide} benchmark dataset
using  \kwarg~\citep{ignatieva2021kwarg},
\argweaver~\citep{rasmussen2014genome,hubisz2020inference},
\tsinfer~\citep{kelleher2019inferring},
and \relate~\citep{speidel2019method}.
% While there is some consistency among the inferred ARGs (such as the samples
% FR-F, WA-F and Af-F being grouped into the same clade), there
% are also considerable differences.
The minimum number
of recombination breakpoints required in the absence of recurrent mutation
is 7 for this dataset~\citep{song2003parsimonious},
and the number of inferred breakpoints
under the input mutation and recombination rates
varies from 2 (\relate) to 17 (\argweaver).
% TODO this needs a bit of rewriting with the rest of the paper in mind
% kwarg and argweaver are inferring eARGs, but we need to be careful
% not to confuse people.
The ARGs inferred by the different methods also have major differences
in how recombination information is represented.
Essentially,
\kwarg\ and \argweaver\ infer eARGs that contain
specific recombination \emph{events} for each breakpoint.
On the other hand, \tsinfer\ and
\relate\ do not infer events but rather
infer gARGs with by differing levels of precision about recombination.
The bottom row of Fig~\ref{fig-inferred-args} shows the extent along
the genome to which graph edges are shared between multiple trees.
We can see, for example, that the \relate\ trees ``share'' less topology
than other methods, but it certainly does not infer a series of
\emph{unrelated} local trees (i.e., with either unlabelled internal
nodes or differently labelled nodes in each tree).
The ARGs inferred by \relate\ and \tsinfer\ both contain substantial
information about node sharing across trees; for example,
both infer that Af-f, Fr-f and Wa-f share a common ancestor along
the entire sequence.
They contain a level of detail about recombination
that lies somewhere between a sequence of unrelated local trees on one
extreme and a completely precise event ARG on the other.

@@

The gARG encoding discussed and concretely defined in this manuscript
leads to highly efficient storage and processing of ARG data,
and has already been in use for several years.
The succinct tree sequence data structure (usually known as a ``tree sequence''
for brevity)
is a concrete encoding of a gARG, as discussed in this article.
It was originally developed as part of the \texttt{msprime}
simulator~\citep{kelleher2016efficient} and has subsequently been
extended and applied to forward-time
simulations~\citep{kelleher2018efficient,haller2018tree},
inference from data~\citep{kelleher2019inferring,wohns2022unified},
and calculation of population genetics statistics~\citep{ralph2020efficiently}.
The succinct tree sequence encoding extends the basic definition
of a gARG provided here by stipulating a
simple tabular representation and nodes and edges,
and also defining a concise and lossless representation of
sequence variation using the ``site'' and  ``mutation'' tables.
(Technically, edges in tree sequence terminology would be better
described as ``edge-intervals'', as each describes a single contiguous
interval of genome inheritance between a pair of nodes. This
denormalisation of the gARG data model in \texttt{tskit} is for efficiency purposes.)
The \texttt{tskit} library is a liberally
licensed open source toolkit that provides a comprehensive suite
of tools for working with gARGs (encoded as a succinct tree sequence).
Based on core functionality written
in C, it provides interfaces in C, Python and Rust.
Tskit is mature software, widely used in population genetics, and
has been incorporated into several downstream
applications~\citep[e.g.,][]{haller2019slim,speidel2019method,
adrion2020community,
terasaki2021geonomics,
baumdicker2021efficient,
fan2022genealogical,korfmann2022weak,
mahmoudi2022bayesian,petr2022slendr,rasmussen2022espalier,
zhang2023biobank,nowbandegani2023extremely,ignatieva2023distribution,
fan2023likelihood}.

@@

% % Do downstream applications actually make use of such full ARGs?
Besides the inherent limitations that exist on inferring fully
precise ARGs from data,
we should also consider the value that such precise estimates provide
for downstream applications.
% To date, few make use such detailed recombination information.
Most applications work by examining local trees independently.
For example, the \relate\ selection test~\citep{speidel2019method}
obtains $p$-values by computing clade size probabilities conditional
on the timing of coalescence events in a given local tree.
In their method
for estimating dispersal rates and the locations of genetic
ancestors,
\cite{osmond2021estimating} downsample trees along the genome
so that they can be regarded as approximately independent.
Similarly, \cite{fan2023likelihood} compute the likelihood
of an ARG under a particular demographic model as the product
over a sample of widely-separated local trees, assumed to be independent.
The SIA method for detecting selection~\citep{hejase2022deep}
encodes local trees as a set of lineage counts at discrete
time intervals, and uses these as feature for a
type of machine learning algorithm
that takes ``temporal'' correlations into account. Thus,
while SIA does use some of the information about local tree correlation,
clearly much of the detail about recombination
events in an ARG is lost.
The \texttt{tsdate} algorithm (and related naive estimator
of ancestral location) uses much more between-tree
information~\citep{wohns2022unified}, explicitly using the gARG
encoding to reason about ancestral nodes.
The main application for fully precise ARGs thus far has been
to compute a likelihood under the coalescent with
recombination~\citep[e.g.][]{kuhner2000maximum,mahmoudi2022bayesian},
which currently requires the details of all recombination
events to be known.

>>>>

\documentclass{article}
\usepackage[letterpaper,top=2cm,bottom=2.5cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{geometry}%
\usepackage{color}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{nameref}
\usepackage[right]{lineno}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{tikz}
\usetikzlibrary{math,calc,positioning}
\usepackage{url}
\usepackage[symbol]{footmisc}
% JK: turning this off for the moment as I keep clicking through on links
% to the bibliography while reading the text and it's intensely annoying.
% Can reinstate when we're ready to preprint
% \usepackage[hidelinks]{hyperref}

@@

\linenumbers
\section{Introduction}
\label{sec-intro}
Estimating the genetic genealogy of a set of genome sequences
under the influence of recombination,
usually known as an Ancestral Recombination Graph (ARG), is a long-standing
goal in genetics.
Broadly speaking, an ARG describes the different paths of genetic inheritance
caused by recombination, and encodes a sequence of correlated genealogical
trees along the genome.
Recent breakthroughs
in large-scale inference
methods~\citep{rasmussen2014genome,kelleher2019inferring,speidel2019method,
schaefer2021ancestral,wohns2022unified,zhang2023biobank,zhan2023towards}
and data representation~\citep{kelleher2016efficient,kelleher2018efficient}
have raised the realistic prospect of ARG-based analysis becoming a standard part
of the population and statistical genetics toolkit~\citep{hejase2020summary}.
Applications using inferred ARGs as input have begun to
appear~\citep{osmond2021estimating,fan2022genealogical,hejase2022deep,zhang2023biobank,
nowbandegani2023extremely,ignatieva2023distribution}
and many more are sure to
follow~\citep{harris2019database,harris2023using}.
See \citet{lewanski2023era} for a biologically oriented introduction to ARGs.

@@

% JK: this para need a major update, it's quite weak right now.
The word ``ARG'' is now widely used to refer to genetic genealogies in
general~\citep[e.g.][]{mathieson2020ancestry,hejase2020summary,
schaefer2021ancestral,harris2023using,zhang2023biobank},
encompassing the varied outputs from modern simulation and
inference methods~\citep{rasmussen2014genome, palamara2016argon, haller2018tree,
kelleher2019inferring, speidel2019method, baumdicker2021efficient, zhang2023biobank}.
This broad usage means there is a lack of consensus on the
exact object being defined, which causes problems, for example,
when evaluating inferred ARGs, and even when defining
what the criteria for success should be.
It also means there is no shared format for interchanging ARG data between
programs, resulting in the assorted negative consequences with which
geneticists are only too familiar~\citep{excoffier2006computer,gower2022demes}.
Perhaps most importantly, definitions linked to
the original stochastic process lead to an
overly narrow interpretation of ARGs, which is significantly
out of step with modern methods and large-scale datasets, and
can lead to missed opportunities in their application and usage.
Such problems go much deeper than pedantic discussion over terminology;
they directly inform what we should be attempting to estimate, how we
assess success, and the expected limitations on the precision
and accuracy of our inferences.
% Thanks for the phrase Gideon and Alex!
If not addressed, these problems will
impede progress in applying genetic genealogies to understand the biological world.

@@

\section{Example inferred ARGs}
\label{sec-example-inferred-args}

\begin{figure} \begin{center}
\includegraphics[width=\textwidth]{illustrations/inference.pdf} \end{center}
\caption{\label{fig-inferred-args} Inference of an 11 sample ARG over a 2.7kb
region of the \textit{Drosophila melanogaster} ADH locus~\citep{kreitman1983nucleotide}.
Results for four different methods based on
(A) parsimony, (B) sampling the SMC$\prime$, (C) copying from
inferred ancestral haplotypes, and (D) correlated local tree estimation.
Edge colours are based on the time of the edge's child node
(lighter: older; darker: younger).
Position on the X and Y axes for the graph visualization is arbitrary.
Line width and node colour are as described in Fig.~\ref{fig-simplification}.
The graphic below each ARG shows the genomic intervals for each graph edge in the
corresponding ARGs: edges are horizontal lines, stacked in time order
(edges with youngest children at the bottom), with vertical dashed lines indicating
breakpoints between local trees.
}
\end{figure}

The scalability gains made by recent ARG inference methods have been, in part, due to inferring
lower levels of precision about recombination. Indeed, a common statement in the literature is
that modern tools such as \tsinfer \citep{kelleher2019inferring} and \relate \citep{speidel2019method}
construct not an ARG but "only" a sequence of local trees \[cite examples from \#38\]. While this statement
is correct inasmuch as neither method attempts to infer the explicit recombination events that cause
adjacent trees to differ, it is certainly not the case that they infer independent local trees.
In fact, both methods take advantage of genome-wide mutation information to construct genealogies,
and as a result nodes and edges persist across multiple local trees.
Their results can therefore be thought of as an inheritance graph, and represented in the gARG format.

To illustrate the varying levels of information that can be captured in a gARG,
Fig.~\ref{fig-inferred-args} presents graph representations of genealogies output by two tools that
\emph{do} explicitly infer recombination events, namely \kwarg \citep{ignatieva2021kwarg} and \argweaver
\citep{rasmussen2014genome}, and two which do not (tsinfer and Relate). All inferences are based on the
same underlying dataset from \citet{kreitman1983nucleotide}, a standard benchmark in the classical
ARG literature, for which the minimum number of recombination breakpoints in the absence of recurrent
mutation is 7~\citep{song2003parsimonious}. This dataset comprises 43 biallelic SNPs spanning 2.4Kb of the
\emph{D.~melanogaster} ADH locus on chromosome 2L. Where required for inference purposes we
assume a mutation rate of \SI{5.49e-9}{site^{-1}generation^{-1}} and a recombination rate of
\SI{8.4e-9}{site^{-1}generation^{-1}} [CITE stdpopsim] and a constant effective population size of 1000,000.
The version numbers of inference software, together with the specific inference parameters
needed to reproduce the ARGs, are available from our GitHub repository.

Note that both \kwarg and \argweaver produce many ARGs
(those shown in Fig.~\ref{fig-inferred-args} A\& B are arbitrarily selected examples).
This stems from the fact that \kwarg searches the space of plausible ARGs,
outputting minimal ones using heuristics, and ARGweaver samples from the posterior distribution over
ARGs given the data and a fixed model (the SMC$\prime$, an approximation of the full coalescent
with recombination). In contrast, \tsinfer and \relate (Fig.~\ref{fig-inferred-args} C \& D) are
deterministic methods, aiming to find a single ``best guess'' ARG.
The ability of a gARG to encode the effects of recombination, rather
than having to precisely specify each event, is a important contributor to this aim,
as most recombination events cannot be uniquely identified from the data.
Indeed, there is considerable variation in the number of recombinations inferred by
different methods, with e.g.~\argweaver (Fig.~\ref{fig-inferred-args} B) suggesting
far more than are required under minimal parsimony assumptions (Fig.~\ref{fig-inferred-args} A).

A rough feel for the amount of recombination in each ARG is given by the distribution
of number of parents per node, as highlighted by the node colour scheme.
In Fig.~\ref{fig-inferred-args} A \& B, each recombination node has exactly 2 parents,
corresponding to a single recombination event with a specific breakpoint. Assuming that each
breakpoint occurs a a different location (the "infinite breakpoints" model), then the
number of breakpoints should exactly equal the number of 2-parent (brown) nodes in the plot.
In contrast, Figure Fig.~\ref{fig-inferred-args} C \& D have ARG nodes with greater
than two parents, indicating deliberate imprecision: these are cases where multiple recombination events
have been ``pushed down'' onto a single more recent node. Note, however, that the number
of breakpoints in these two ARGs is smaller than the number of ARG nodes with multiple parents,
because those nodes frequently share breakpoint positions. The sharing of breakpoint locations
between multiple-parent nodes indicates that in \tsinfer and \relate ARGs it may be
difficult to condense each transition between trees into a set of simple SPR moves.

Shading by coalescent span reveals that many non-recombinant (green) nodes are
``locally unary'' (see section \ref{sec-locally-unary-edges}). In other words,
in some regions of the genome, these nodes are present in the ancestry but do
not correspond to coalescence. As argued previously, this can provide potentially useful
information about shared ancestry. Note however, that such unary segments are not inferred
in the \relate ARG: other than the samples it has entirely solid nodes.
Only the \tsinfer results have nodes with more than two children (polytomies);
in Appendix \ref{sec-cell-lineages-and-args} we argue that this reflects
uncertainty in the order of coalescent events, in an equivalent
manner to the lack of precision in recombination nodes with more
than two parents. The clearest example is the polytomy involving the nodes
\noderef{Fr-F}, \noderef{Wa-F}, and \noderef{Af-F}, which happen to
have identical sequences. \kwarg, \argweaver, and \relate require
bifurcating trees, so are forced
to pick an arbitrary order of branching
(hence Fig.~\ref{fig-inferred-args} A and Fig.~\ref{fig-inferred-args} B disagree on the
relationship between these samples). TsThe number of nodes in Fig.~\ref{fig-inferred-args} C
that have more than two children also hints at considerable uncertainty in
the order of branching elsewhere in the ARG.
% TODO add explanation of the bottom row.
% https://github.com/tskit-dev/what-is-an-arg-paper/issues/386
%The example of successive simplification of a simulated ARG
%in Fig.~\ref{fig-simplification} illustrates the point
%that is is possible to \emph{store} arbitrary levels of precision
%about recombination in a gARG.
%It is also possible to \emph{infer} various levels of precision about
%recombination.
%To illustrate this, and to demonstrate the structural
%heterogeneity of the ARGs inferred  by current methods,
%Fig.~\ref{fig-inferred-args} shows the results of ARG inference
% To illustrate these points and to emphasise the heterogeneity
% of the ARGs inferred by current methods, we inferred ARGs for the
%for the classical \citet{kreitman1983nucleotide} benchmark dataset
%using  \kwarg~\citep{ignatieva2021kwarg},
%\argweaver~\citep{rasmussen2014genome,hubisz2020inference},
%\tsinfer~\citep{kelleher2019inferring},
%and \relate~\citep{speidel2019method}.
% While there is some consistency among the inferred ARGs (such as the samples
% FR-F, WA-F and Af-F being grouped into the same clade), there
% are also considerable differences.
%The minimum number
%of recombination breakpoints required in the absence of recurrent mutation
%is 7 for this dataset~\citep{song2003parsimonious},
%and the number of inferred breakpoints
%under the input mutation and recombination rates
%varies from 2 (\relate) to 17 (\argweaver).
% TODO this needs a bit of rewriting with the rest of the paper in mind
% kwarg and argweaver are inferring eARGs, but we need to be careful
% not to confuse people.
%The ARGs inferred by the different methods also have major differences
%in how recombination information is represented.
%Essentially,
%\kwarg\ and \argweaver\ infer eARGs that contain
%specific recombination \emph{events} for each breakpoint.
%On the other hand, \tsinfer\ and
%\relate\ do not infer events but rather
%infer gARGs with by differing levels of precision about recombination.

The bottom row of Fig~\ref{fig-inferred-args} shows the extent along
the genome to which graph edges are shared between multiple trees.
We can see, for example, that the \relate\ trees ``share'' less topology
than other methods, but it certainly does not infer a series of
\emph{unrelated} local trees
%(i.e., with either unlabelled internal
%nodes or differently labelled nodes in each tree).
The ARGs inferred by \relate\ and \tsinfer\ both contain substantial
information about node sharing across trees; for example,
both infer that Af-f, Fr-f and Wa-f share a common ancestor along
the entire sequence.
They contain a level of detail about recombination
that lies somewhere between a sequence of unrelated local trees on one
extreme and a completely precise event ARG on the other.

@@

The gARG encoding discussed and concretely defined in this manuscript
leads to highly efficient storage and processing of ARG data,
and has already been in use for several years.
The succinct tree sequence data structure (usually known as a ``tree sequence''
for brevity)
is a concrete encoding of a gARG, as discussed in this article.
It was originally developed as part of the \texttt{msprime}
simulator~\citep{kelleher2016efficient} and has subsequently been
extended and applied to forward-time
simulations~\citep{kelleher2018efficient,haller2018tree},
inference from data~\citep{kelleher2019inferring,wohns2022unified},
and calculation of population genetics statistics~\citep{ralph2020efficiently}.
The succinct tree sequence encoding extends the basic definition
of a gARG provided here by stipulating a
simple tabular representation and nodes and edges,
and also defining a concise and lossless representation of
sequence variation using the ``site'' and  ``mutation'' tables.
(Technically, edges in tree sequence terminology would be better
described as ``edge-intervals'', as each describes a single contiguous
interval of genome inheritance between a pair of nodes. This
denormalisation of the gARG data model in \texttt{tskit} is for efficiency purposes.)
The \texttt{tskit} library is a liberally
licensed open source toolkit that provides a comprehensive suite
of tools for working with gARGs (encoded as a succinct tree sequence).
Based on core functionality written
in C, it provides interfaces in C, Python and Rust.
Tskit is mature software, widely used in population genetics, and
has been incorporated into several downstream
applications~\citep[e.g.,][]{haller2019slim,speidel2019method,
adrion2020community,
terasaki2021geonomics,
baumdicker2021efficient,
fan2022genealogical,korfmann2022weak,
mahmoudi2022bayesian,petr2022slendr,rasmussen2022espalier,
zhang2023biobank,nowbandegani2023extremely,ignatieva2023distribution}.

@@

% % Do downstream applications actually make use of such full ARGs?
Besides the inherent limitations that exist on inferring fully
precise ARGs from data,
we should also consider the value that such precise estimates provide
for downstream applications.
To date, few downstream applications using estimated ARGs
make use of complete recombination information.
Most applications work by examining local trees independently.
For example, the \relate\ selection test~\citep{speidel2019method}
obtains $p$-values by computing clade size probabilities conditional on the timing of coalescence events in a given local tree.
In their method
for estimating dispersal rates and the locations of genetic
ancestors,
\cite{osmond2021estimating} downsample trees along the genome
so that they can be regarded as approximately independent.
The SIA method for detecting selection~\citep{hejase2022deep}
encodes local trees as a set of lineage counts at discrete
time intervals, and uses these as feature for a
type of machine learning algorithm
that takes ``temporal'' correlations into account. Thus,
while SIA does use some of the information about local tree correlation,
clearly much of the detail about recombination
events in an ARG is lost.
The \texttt{tsdate} algorithm (and related naive estimator
of ancestral location) uses much more between-tree
information~\citep{wohns2022unified}, explicitly using the gARG
encoding to reason about ancestral nodes.
The main application for fully precise ARGs thus far has been
to compute a likelihood under the coalescent with
recombination~\citep[e.g.][]{kuhner2000maximum,mahmoudi2022bayesian},
which currently requires the details of all recombination
events to be known.
